---
title: "Assignment 8"
author: "Megan Willis-Jackson, Claire Wang, Ignacio Lafuente"
date: "3/25/2022"
output: html_document
---

# Load Libraries
```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(sf)
library(survey)
library(srvyr)
library(od)
library(ggspatial)
library(scenRios)
```

# Load trip generation data and skims
```{r}
zones <- here("existing",
              "data",
              "existing_sanjose.csv") %>%
  read_csv(show_col_types = FALSE)

skims <- here("existing",
              "data",
              "skims.csv") %>%
  read_csv(show_col_types = FALSE)

```

# Calculate average travel time by trip purpose

```{r}
#download and temporarily store NHTS data

temp <- tempfile()
download.file("https://nhts.ornl.gov/assets/2016/download/csv.zip", temp)

trips <- read_csv(unz(temp, "trippub.csv"), 
                      show_col_types = FALSE) %>%
  filter(HH_CBSA == "41940")

unlink(temp)



```

```{r}
#trip purpose variable

trips <- trips %>%
  mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
                                WHYTO == "02" ~ TRUE,
                                WHYFROM == "01" ~ TRUE,
                                WHYFROM == "02" ~ TRUE,
                                TRUE ~ FALSE)) %>%
  mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
  mutate(purpose = case_when(home_based & work ~ "HBW",
                            home_based ~ "HBO",
                            TRUE ~ "NHB"))


#create survey object, calculate avg travel time by trip purpose

trips_svy <- trips %>%
  as_survey(weights = WTTRDFIN)

ttime_by_purpose <- trips_svy %>%
  group_by(purpose) %>%
  summarise(avg_time = survey_mean(TRVLCMIN))

ttime_by_purpose


```

# Calculate minimum travel time across all modes

```{r}
skims <- skims %>%
  mutate(min_time = pmin(transit_time, 
                         car_time,
                         bike_time,
                         walk_time,
                         na.rm = TRUE)) 

```

# Friction Factors

## Gamma Function

```{r}
skims <- skims %>%
  mutate(F_HBW = min_time^-0.503*exp(-0.078*min_time),
         F_HBO = min_time^-3.993*exp(-0.019*min_time),
         F_NHB = min_time^-3.345*exp(-0.003*min_time)) 


```

# Estimate travel flows
Only run the travel flows code if we are changing the 
Friction Factor type, because it takes so long to 
run. The lists are saved and can be loaded directly
using the code directly below.

```{r}
#Load travel flows
#HBO
HBO_dist <- readRDS(here("existing",
                         "data",
                         "HBO_dist.Rds"))

#HBW
HBW_dist <- readRDS(here("existing",
                         "data",
                         "HBW_dist.Rds"))

#NHB
NHB_dist <- readRDS(here("existing",
                         "data",
                         "NHB_dist.Rds"))




```

# following chunk not necessary unless changing friction factors
```{r}
#HBO travel flows
HBO_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbo_prod",
                            zone_d = "hbo_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBO",
                            tolerance = 5,
                            max_iter = 50000)

saveRDS(HBO_dist, here("existing",
                       "data",
                       "HBO_dist.Rds"))


#HBW travel flows
HBW_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbw_prod",
                            zone_d = "hbw_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBW",
                            tolerance = 5,
                            max_iter = 50000)

table(HBW_dist$flows$flow > 0)

saveRDS(HBW_dist, here("existing",
                       "data",
                       "HBW_dist.Rds"))


#NHB travel flows
NHB_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "nhb_prod",
                            zone_d = "nhb_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_NHB",
                            tolerance = 5,
                            max_iter = 50000)

table(NHB_dist$flows$flow > 0)

saveRDS(NHB_dist, here("existing",
                       "data",
                       "NHB_dist.Rds"))



```

# Map desire lines

```{r}
zone_boundaries <- here("zones",
              "zones_sanjose.geojson") %>%
  st_read(quiet = TRUE)

desire_lines_HBO <- od_to_sf(HBO_dist$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

desire_lines_HBW <- od_to_sf(HBW_dist$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

desire_lines_NHB <- od_to_sf(NHB_dist$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)



#very messy plot with little detail available
ggplot(desire_lines_HBO) +
  annotation_map_tile(type = "cartolight", zoom = 10, progress = "none") +
  geom_sf(aes(alpha = flow)) +
  theme_void()

#most attractive HBO zone
HBO_big_attraction <- zones[zones$hbo_attr_bal == max(zones$hbo_attr_bal),]$GEOID

HBO_big_attraction

#most attractive HBW zone
HBW_big_attraction <- zones[zones$hbw_attr_bal == max(zones$hbw_attr_bal),]$GEOID

HBW_big_attraction

#most attractive HBW zone
NHB_big_attraction <- zones[zones$nhb_attr_bal == max(zones$nhb_attr_bal),]$GEOID

NHB_big_attraction

# the same zone is the largest attraction for 
#all three trip types: 06085505006

desire_lines_one_zone_HBO <- desire_lines_HBO %>%
  filter(d_id == big_attraction)

# pull out the x/y coordinates of the desire lines
# to enable curving them on the map, just for aesthetics
desire_lines_coord_HBO <- desire_lines_one_zone_HBO %>%
  st_cast("POINT") %>%
  st_coordinates() %>%
  .[c(TRUE,FALSE),] %>%
  cbind(desire_lines_one_zone_HBO %>%
          st_cast("POINT") %>%
          st_coordinates() %>%
          .[!c(TRUE,FALSE),]) %>%
  cbind(desire_lines_one_zone_HBO)

ggplot(desire_lines_one_zone_HBO) +
  annotation_map_tile(type = "cartolight", zoom = 11, progress = "none") +
  geom_sf(color = NA) +
  geom_curve(data = desire_lines_coord_HBO,
             aes(x = X, y = Y, xend = X.1, yend = Y.1, alpha = flow), curvature = .2) +
  theme_void()



```



