---
title: "Assignment02"
author: "Megan Willis-Jackson, Claire Wang, Ignacio Lafuente"
date: "2/3/2022"
output: html_document
---

```{r libraries, echo=F, message=F, warning=F}
library(tidyverse)
library(tidycensus)
library(ggplot2)
library(sf)
library(units)
library(tigris)


```

# Load Variables

```{r}
sanjose_msa <- core_based_statistical_areas() %>%
  filter(GEOID == "41940")

# view(load_variables(2019, "acs5"))

vars <- c(med_income = "B06011_001",
          total_hhs = 'B08203_001',
          no_veh_hh = 'B08203_002',
          one_veh_hh = 'B08201_003',
          two_veh_hh = 'B08201_004',
          three_veh_hh = 'B08201_005',
          fourmore_veh_hh = 'B08201_006',
          no_worker_hh = 'B08202_002',
          one_worker_hh = 'B08202_003',
          two_worker_hh = 'B08202_004',
          threemore_worker_hh = 'B08202_005',
          total_worker = 'B08006_001',
          total_wfh = 'B08006_017',
          hh_1person = 'B08201_007',
          hh_2person = 'B08201_013',
          hh_3person = 'B08201_019',
          hh_4person_plus = 'B08201_025',
          inc_lt_10k = 'B19001_002',
          inc_btw_10k_15k = 'B19001_003',
          inc_btw_15k_20k = 'B19001_004',
          inc_btw_20k_25k = 'B19001_005',
          inc_btw_25k_30k = 'B19001_006',
          inc_btw_30k_35k = 'B19001_007',
          inc_btw_35k_40k = 'B19001_008',
          inc_btw_40k_45k = 'B19001_009',
          inc_btw_45k_50k = 'B19001_010',
          inc_btw_50k_60k = 'B19001_011',
          inc_btw_60k_75k = 'B19001_012',
          inc_btw_75k_100k = 'B19001_013',
          inc_btw_100k_125k = 'B19001_014',
          inc_btw_125k_150k = 'B19001_015',
          inc_btw_150k_200k = 'B19001_016',
          inc_gt_200k = 'B19001_017') 

census <- get_acs(geography = "tract",
                  state = "CA",
                  variables = vars,
                  output = "wide",
                  geometry = T) %>%
  st_filter(sanjose_msa)

tracts <- get_acs(geography = "tract",
                  state = "CA",
                  variables = "B06011_001",
                  output = "wide",
                  geometry = T) %>%
  st_filter(sanjose_msa)



ggplot(census) +
  geom_sf(aes(fill = med_incomeE), color = alpha("white", .1)) +
  theme_void()

#download employment data
#NOTE: I think we also will need the LEHD residence area characteristics,
#to determine which RESIDENTS will no longer have to commute, in addition to 
#which jobs in the MSA will no longer have commuters.

#creates new variables, including a WFH variable which
#assumes that the job is able to be at least partially remote.
lehd_blocks_jobs <- read_csv("https://lehd.ces.census.gov/data/lodes/LODES7/ca/wac/ca_wac_S000_JT00_2019.csv.gz", 
                             show_col_types = F) %>%
  rename(total_jobs = C000) %>%
  mutate(basic_jobs = CNS01+CNS02+CNS03+CNS04+CNS05+CNS06+CNS08+CNS09) %>%
  rename(retail_jobs = CNS07) %>%
  mutate(service_jobs = total_jobs - basic_jobs - retail_jobs,
         WFH_jobs = CNS09+CNS10+CNS12+CNS13+CNS19+CNS20) %>%
  select(w_geocode, total_jobs, basic_jobs, retail_jobs, service_jobs, WFH_jobs)

#load the residence area characteristics
lehd_blocks_emp <- read_csv("https://lehd.ces.census.gov/data/lodes/LODES7/ca/rac/ca_rac_S000_JT00_2019.csv.gz",
                            show_col_types = F) %>%
  rename(total_emp = C000) %>%
  mutate(basic_emp = CNS01+CNS02+CNS03+CNS04+CNS05+CNS06+CNS08+CNS09) %>%
  rename(retail_emp = CNS07) %>%
  mutate(service_emp = total_emp - basic_emp - retail_emp,
         WFH_emp = CNS09+CNS10+CNS12+CNS13+CNS19+CNS20) %>%
  select(h_geocode, total_emp, basic_emp, retail_emp, service_emp, WFH_emp)


#collapse lehd variables to census tracts
lehd_tracts_jobs <- lehd_blocks_jobs %>%
  mutate(w_geocode = as.character(w_geocode)) %>%
  mutate(GEOID = substr(w_geocode, 1, 11)) %>%
  select(-w_geocode) %>%
  group_by(GEOID) %>%
  summarise(across(everything(), ~sum(.))) 

lehd_tracts_emp <- lehd_blocks_emp %>%
  mutate(h_geocode = as.character(h_geocode)) %>%
  mutate(GEOID = substr(h_geocode, 1, 11)) %>%
  select(-h_geocode) %>%
  group_by(GEOID) %>%
  summarise(across(everything(), ~sum(.))) 

census <- census %>%
  left_join(lehd_tracts_jobs, by = "GEOID") %>%
  left_join(lehd_tracts_emp, by = "GEOID") %>%
  select(-contains("M", ignore.case = F))


```

# Create Alternative Dataset

The main assumption here is that the industries that we have classified
as "WFH" will experience a 40% decrease in work days commuting from the 
baseline, i.e. from 5 days per week to 3 days per week.

- generate assumption of % days work from home (e.g. 40%)
- generate 40% wfh-able jobs and employees
- generate: total workers - WFH - 40% * (WFH-able jobs + WFH-able employees)
40% is an assumption
- assume people WFH is not counted in the total jobs
- assume that there is negligible overlap between people working in a tract and living 
in that tract. (based on random sampling of LEHD data)

```{r}
census_alt <- census %>%
  mutate(wfh_percent = .4,
         alt_wfh_emp = round(WFH_emp * (1-wfh_percent), digits = 0),
         alt_wfh_jobs = round(WFH_jobs * (1-wfh_percent), digits = 0),
         alt_total_jobs = round(total_jobs - wfh_percent * WFH_jobs, digits = 0),
         alt_total_emp = round(total_emp * (1 - total_wfhE/total_workerE) - 
                                 wfh_percent * WFH_emp), digits = 0)
  




```

# Save Datasets

```{r}
existing <- write_csv(census, "existing/existing_sanjose.csv", append = F)

zones_sanjose <- st_write(tracts, "zones/zones_sanjose2.geojson", append = F)

alternative <- write_csv(census_alt, "alternative/alternative_sanjose.csv", append = F)




```









