---
title: "Assignment 8"
author: "Megan Willis-Jackson, Claire Wang, Ignacio Lafuente"
date: "3/25/2022"
output: html_document
---

# Load Libraries
```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(sf)
library(survey)
library(srvyr)
library(od)
library(ggspatial)
library(scenRios)
```


# EXISTING CONDITIONS

# Load trip generation data and skims
```{r}
zones <- here("existing",
              "data",
              "existing_sanjose.csv") %>%
  read_csv(show_col_types = FALSE)

skims <- here("existing",
              "data",
              "skims.csv") %>%
  read_csv(show_col_types = FALSE)

```

# Calculate average travel time by trip purpose

```{r}
#download and temporarily store NHTS data

temp <- tempfile()
download.file("https://nhts.ornl.gov/assets/2016/download/csv.zip", temp)

trips <- read_csv(unz(temp, "trippub.csv"), 
                      show_col_types = FALSE) %>%
  filter(HH_CBSA == "41940")

unlink(temp)



```

```{r}
#trip purpose variable

trips <- trips %>%
  mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
                                WHYTO == "02" ~ TRUE,
                                WHYFROM == "01" ~ TRUE,
                                WHYFROM == "02" ~ TRUE,
                                TRUE ~ FALSE)) %>%
  mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
  mutate(purpose = case_when(home_based & work ~ "HBW",
                            home_based ~ "HBO",
                            TRUE ~ "NHB"))


#create survey object, calculate avg travel time by trip purpose

trips_svy <- trips %>%
  as_survey(weights = WTTRDFIN)

ttime_by_purpose <- trips_svy %>%
  group_by(purpose) %>%
  summarise(avg_time = survey_mean(TRVLCMIN))

ttime_by_purpose


```

# Calculate minimum travel time across all modes

```{r}
skims <- skims %>%
  mutate(min_time = pmin(transit_time, 
                         car_time,
                         bike_time,
                         walk_time,
                         na.rm = TRUE)) 

```

# Friction Factors

## Gamma Function

```{r}
#Large MPO 1
# skims <- skims %>%
#   mutate(F_HBW = min_time^-0.503*exp(-0.078*min_time),
#          F_HBO = min_time^-3.993*exp(-0.019*min_time),
#          F_NHB = min_time^-3.345*exp(-0.003*min_time))


#Large MPO 2
# skims <- skims %>%
#   mutate(F_HBW = min_time^-1.65*exp(-0.0398*min_time),
#          F_HBO = min_time^-1.51*exp(-0.018*min_time),
#          F_NHB = min_time^-1.94*exp(-0.116*min_time))


```

## Power function
We tried a gamma function for both Large MPO 1 and 2, and
the best fit ended up being the power function. It may be 
possible to get average travel time closer to the average
indicated by the NHTS data, but due to time constraints,
we will make due with the power function.

```{r}

skims <- skims %>%
  mutate(F_HBO = min_time^-2,
         F_HBW = min_time^-2,
         F_NHB = min_time^-2) 



```

# Estimate travel flows
Only run the travel flows code if we are changing the 
Friction Factor type, because it takes so long to 
run. The lists are saved and can be loaded directly
using the code directly below.

```{r}
#Load travel flows
#HBO
HBO_dist <- readRDS(here("existing",
                         "data",
                         "HBO_dist_power.Rds"))

#HBW
HBW_dist <- readRDS(here("existing",
                         "data",
                         "HBW_dist_power.Rds"))

#NHB
NHB_dist <- readRDS(here("existing",
                         "data",
                         "NHB_dist_power.Rds"))




```

# following chunk not necessary unless changing friction factors

We initially ran the model using gamma function parameters for Large MPO 1, but the 
average travel times this yielded were about half as long as what is suggested by the 
NHTS data. Looking at the 


```{r}
#HBO travel flows
HBO_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbo_prod",
                            zone_d = "hbo_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBO",
                            tolerance = 5,
                            max_iter = 25000)

saveRDS(HBO_dist, here("existing",
                       "data",
                       "HBO_dist_power.Rds"))


#HBW travel flows
HBW_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "hbw_prod",
                            zone_d = "hbw_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBW",
                            tolerance = 5,
                            max_iter = 25000)

table(HBW_dist$flows$flow > 0)

saveRDS(HBW_dist, here("existing",
                       "data",
                       "HBW_dist_power.Rds"))


#NHB travel flows
NHB_dist <- grvty_balancing(od_zones = zones,
                            friction = skims,
                            zone_id = "GEOID",
                            zone_o = "nhb_prod",
                            zone_d = "nhb_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_NHB",
                            tolerance = 5,
                            max_iter = 25000)

table(NHB_dist$flows$flow > 0)

saveRDS(NHB_dist, here("existing",
                       "data",
                       "NHB_dist.Rds"))



```


# Compare caluclated times to observed times

```{r}


#HBO existing

HBO_flows <- HBO_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBO_flow = flow)

skims <- skims %>%
  left_join(HBO_flows, by = c("fromId", "toId")) %>%
  replace_na(list(HBO_flow = 0)) %>%
  mutate(HBO_total_time = HBO_flow * min_time)

HBO_mean_time <- sum(skims$HBO_total_time) / sum(skims$HBO_flow)



#HBW existing

HBW_flows <- HBW_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         HBW_flow = flow)

skims <- skims %>%
  left_join(HBW_flows, by = c("fromId", "toId")) %>%
  replace_na(list(HBW_flow = 0)) %>%
  mutate(HBW_total_time = HBW_flow * min_time)

HBW_mean_time <- sum(skims$HBW_total_time) / sum(skims$HBW_flow)



#NHB existing

NHB_flows <- NHB_dist$flows %>%
  rename(fromId = o_id,
         toId = d_id,
         NHB_flow = flow)

skims <- skims %>%
  left_join(NHB_flows, by = c("fromId", "toId")) %>%
  replace_na(list(NHB_flow = 0)) %>%
  mutate(NHB_total_time = NHB_flow * min_time)

NHB_mean_time <- sum(skims$NHB_total_time) / sum(skims$NHB_flow)




```

## Check averages

```{r}
ttime_by_purpose

HBO_mean_time

HBW_mean_time

NHB_mean_time


```




# Map desire lines

```{r}
zone_boundaries <- here("zones",
              "zones_sanjose.geojson") %>%
  st_read(quiet = TRUE)

desire_lines_HBO <- od_to_sf(HBO_dist$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

desire_lines_HBW <- od_to_sf(HBW_dist$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

desire_lines_NHB <- od_to_sf(NHB_dist$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)




#most attractive HBO zone
HBO_big_attraction <- zones[zones$hbo_attr_bal == max(zones$hbo_attr_bal),]$GEOID

HBO_big_attraction

#most attractive HBW zone
HBW_big_attraction <- zones[zones$hbw_attr_bal == max(zones$hbw_attr_bal),]$GEOID

HBW_big_attraction

#most attractive HBW zone
NHB_big_attraction <- zones[zones$nhb_attr_bal == max(zones$nhb_attr_bal),]$GEOID

NHB_big_attraction

# the same zone is the largest attraction for 
#all three trip types: 06085505006

desire_lines_one_zone_HBO <- desire_lines_HBO %>%
  filter(d_id == HBO_big_attraction)

desire_lines_one_zone_HBW <- desire_lines_HBW %>%
  filter(d_id == HBW_big_attraction)

desire_lines_one_zone_NHB <- desire_lines_NHB %>%
  filter(d_id == NHB_big_attraction)

# combine one zone desire lines for all trip types into one dataset
desire_lines_one_zone_all <- desire_lines_one_zone_HBO %>%
  rename(hbo_flow = flow) %>%
  as.data.frame() %>%
  full_join(desire_lines_one_zone_HBW %>%
              rename(hbw_flow = flow) %>%
              as.data.frame(), 
            by = c("o_id", "d_id", "geometry")) %>%
  full_join(desire_lines_one_zone_NHB %>%
              rename(nhb_flow = flow) %>%
              as.data.frame(),
            by = c("o_id", "d_id", "geometry")) %>%
  st_as_sf()


# pull out the x/y coordinates of the desire lines
# to enable curving them on the map, just for aesthetics
# HBO
desire_lines_coord_HBO <- desire_lines_one_zone_HBO %>%
  st_cast("POINT") %>%
  st_coordinates() %>%
  .[c(TRUE,FALSE),] %>%
  cbind(desire_lines_one_zone_HBO %>%
          st_cast("POINT") %>%
          st_coordinates() %>%
          .[!c(TRUE,FALSE),]) %>%
  cbind(desire_lines_one_zone_HBO)

# HBW
desire_lines_coord_HBW <- desire_lines_one_zone_HBW %>%
  st_cast("POINT") %>%
  st_coordinates() %>%
  .[c(TRUE,FALSE),] %>%
  cbind(desire_lines_one_zone_HBW %>%
          st_cast("POINT") %>%
          st_coordinates() %>%
          .[!c(TRUE,FALSE),]) %>%
  cbind(desire_lines_one_zone_HBW)

# NHB
desire_lines_coord_NHB <- desire_lines_one_zone_NHB %>%
  st_cast("POINT") %>%
  st_coordinates() %>%
  .[c(TRUE,FALSE),] %>%
  cbind(desire_lines_one_zone_NHB %>%
          st_cast("POINT") %>%
          st_coordinates() %>%
          .[!c(TRUE,FALSE),]) %>%
  cbind(desire_lines_one_zone_NHB)

# All
desire_lines_coord_all <- desire_lines_one_zone_all %>%
  st_cast("POINT") %>%
  st_coordinates() %>%
  .[c(TRUE,FALSE),] %>%
  cbind(desire_lines_one_zone_all %>%
          st_cast("POINT") %>%
          st_coordinates() %>%
          .[!c(TRUE,FALSE),]) %>%
  cbind(desire_lines_one_zone_all)
```

## Plots

```{r}



ggplot(desire_lines_one_zone_HBO) +
  # annotation_map_tile(type = "cartolight", zoom = 11, progress = "none") +
  geom_sf(data = zone_boundaries, fill = "black", color = alpha("white", .1), size = .1) +
  geom_sf(color = NA) +
  geom_curve(data = desire_lines_coord_HBO,
             aes(x = X, y = Y, xend = X.1, yend = Y.1, 
                 alpha = flow,
                 color = flow), curvature = .2, show.legend = F) +
  scale_color_viridis_b(option = "C") +
  theme_void()

#zoom in on downtown for more clarity
left_side <- st_bbox(zone_boundaries %>% filter(GEOID == HBO_big_attraction))$xmin - .01
right_side <- st_bbox(zone_boundaries %>% filter(GEOID == HBO_big_attraction))$xmax + .01
top_side <- st_bbox(zone_boundaries %>% filter(GEOID == HBO_big_attraction))$ymax
bottom_side <- st_bbox(zone_boundaries %>% filter(GEOID == HBO_big_attraction))$ymin

ggplot(desire_lines_one_zone_HBO) +
  geom_sf(data = zone_boundaries, fill = "black", color = alpha("white", .1), size = .1) +
  geom_sf(color = NA) +
  geom_curve(data = desire_lines_coord_HBO,
             aes(x = X, y = Y, xend = X.1, yend = Y.1, 
                 alpha = flow,
                 color = flow), curvature = .2) +
  coord_sf(xlim = c(left_side, right_side),
           ylim = c(bottom_side, top_side)) +
  scale_color_viridis_b(option = "C",
                        name = "Trip Flow",
                        labels = c("3,000","6,000","9,000")) +
  scale_alpha(guide = 'none') +
  theme_void()




```







# ALTERNATIVE CONDITIONS

# Load trip generation data and skims
```{r}
zones_alt <- here("alternative",
                  "data",
                  "alternative_sanjose.csv") %>%
  read_csv(show_col_types = FALSE)

skims_alt <- here("alternative",
                  "data",
                  "skims.csv") %>%
  read_csv(show_col_types = FALSE)

```

# Calculate average travel time by trip purpose

```{r}
#download and temporarily store NHTS data

temp <- tempfile()
download.file("https://nhts.ornl.gov/assets/2016/download/csv.zip", temp)

trips <- read_csv(unz(temp, "trippub.csv"), 
                      show_col_types = FALSE) %>%
  filter(HH_CBSA == "41940")

unlink(temp)



```

```{r}
#trip purpose variable

trips <- trips %>%
  mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
                                WHYTO == "02" ~ TRUE,
                                WHYFROM == "01" ~ TRUE,
                                WHYFROM == "02" ~ TRUE,
                                TRUE ~ FALSE)) %>%
  mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
  mutate(purpose = case_when(home_based & work ~ "HBW",
                            home_based ~ "HBO",
                            TRUE ~ "NHB"))


#create survey object, calculate avg travel time by trip purpose

trips_svy <- trips %>%
  as_survey(weights = WTTRDFIN)

ttime_by_purpose <- trips_svy %>%
  group_by(purpose) %>%
  summarise(avg_time = survey_mean(TRVLCMIN))

ttime_by_purpose


```

# Calculate minimum travel time across all modes

```{r}
skims_alt <- skims_alt %>%
  mutate(min_time = pmin(transit_time, 
                         car_time,
                         bike_time,
                         walk_time,
                         na.rm = TRUE)) 

```

# Friction Factors

## Gamma Function

```{r}
# skims_alt <- skims_alt %>%
#   mutate(F_HBW = min_time^-0.503*exp(-0.078*min_time),
#          F_HBO = min_time^-3.993*exp(-0.019*min_time),
#          F_NHB = min_time^-3.345*exp(-0.003*min_time)) 


```


## Power function
We tried a gamma function for both Large MPO 1 and 2, and
the best fit ended up being the power function. It may be 
possible to get average travel time closer to the average
indicated by the NHTS data, but due to time constraints,
we will make due with the power function.

```{r}

skims_alt <- skims_alt %>%
  mutate(F_HBO = min_time^-2,
         F_HBW = min_time^-2,
         F_NHB = min_time^-2) 



```


# Estimate travel flows
Only run the travel flows code if we are changing the 
Friction Factor type, because it takes so long to 
run. The lists are saved and can be loaded directly
using the code directly below.

```{r}
#Load travel flows
#HBO
HBO_dist_alt <- readRDS(here("alternative",
                             "data",
                             "HBO_dist_alt_power.Rds"))

#HBW
HBW_dist_alt <- readRDS(here("alternative",
                             "data",
                             "HBW_dist_alt_power.Rds"))

#NHB
NHB_dist_alt <- readRDS(here("alternative",
                             "data",
                             "NHB_dist_alt_power.Rds"))




```

# following chunk not necessary unless changing friction factors
```{r}
#HBO travel flows
HBO_dist_alt <- grvty_balancing(od_zones = zones_alt,
                            friction = skims_alt,
                            zone_id = "GEOID",
                            zone_o = "hbo_prod",
                            zone_d = "hbo_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBO",
                            tolerance = 5,
                            max_iter = 25000)

saveRDS(HBO_dist_alt, here("alternative",
                       "data",
                       "HBO_dist_alt_power.Rds"))


#HBW travel flows
HBW_dist_alt <- grvty_balancing(od_zones = zones_alt,
                            friction = skims_alt,
                            zone_id = "GEOID",
                            zone_o = "hbw_prod",
                            zone_d = "hbw_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_HBW",
                            tolerance = 5,
                            max_iter = 25000)

table(HBW_dist$flows$flow > 0)

saveRDS(HBW_dist_alt, here("alternative",
                       "data",
                       "HBW_dist_alt_power.Rds"))


#NHB travel flows
NHB_dist_alt <- grvty_balancing(od_zones = zones_alt,
                            friction = skims_alt,
                            zone_id = "GEOID",
                            zone_o = "nhb_prod",
                            zone_d = "nhb_attr_bal",
                            friction_o_id = "fromId",
                            friction_d_id = "toId",
                            friction_factor = "F_NHB",
                            tolerance = 5,
                            max_iter = 25000)

table(NHB_dist$flows$flow > 0)

saveRDS(NHB_dist_alt, here("alternative",
                       "data",
                       "NHB_dist_alt_power.Rds"))



```

# Map desire lines

```{r}

desire_lines_HBO_alt <- od_to_sf(HBO_dist_alt$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

desire_lines_HBW_alt <- od_to_sf(HBW_dist_alt$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)

desire_lines_NHB_alt <- od_to_sf(NHB_dist_alt$flows, zone_boundaries, silent = TRUE) %>%
  filter(flow > 0)



#most attractive HBO zone
HBO_big_attraction_alt <- zones_alt[zones_alt$hbo_attr_bal == max(zones_alt$hbo_attr_bal),]$GEOID

HBO_big_attraction_alt

#most attractive HBW zone
HBW_big_attraction_alt <- zones_alt[zones_alt$hbw_attr_bal == max(zones_alt$hbw_attr_bal),]$GEOID

HBW_big_attraction_alt

#most attractive HBW zone
NHB_big_attraction_alt <- zones_alt[zones_alt$nhb_attr_bal == max(zones_alt$nhb_attr_bal),]$GEOID

NHB_big_attraction_alt

# the same zone is the largest attraction for 
#all three trip types: 06085505006

desire_lines_one_zone_HBO_alt <- desire_lines_HBO_alt %>%
  filter(d_id == HBO_big_attraction_alt)

desire_lines_one_zone_HBW_alt <- desire_lines_HBW_alt %>%
  filter(d_id == HBW_big_attraction_alt)

desire_lines_one_zone_NHB_alt <- desire_lines_NHB_alt %>%
  filter(d_id == NHB_big_attraction_alt)

# combine one zone desire lines for all trip types into one dataset
desire_lines_one_zone_all_alt <- desire_lines_one_zone_HBO_alt %>%
  rename(hbo_flow_alt = flow) %>%
  as.data.frame() %>%
  full_join(desire_lines_one_zone_HBW_alt %>%
              rename(hbw_flow_alt = flow) %>%
              as.data.frame(), 
            by = c("o_id", "d_id", "geometry")) %>%
  full_join(desire_lines_one_zone_NHB_alt %>%
              rename(nhb_flow_alt = flow) %>%
              as.data.frame(),
            by = c("o_id", "d_id", "geometry")) %>%
  st_as_sf()



# pull out the x/y coordinates of the desire lines
# to enable curving them on the map, just for aesthetics
# HBO
desire_lines_coord_HBO_alt <- desire_lines_one_zone_HBO_alt %>%
  st_cast("POINT") %>%
  st_coordinates() %>%
  .[c(TRUE,FALSE),] %>%
  cbind(desire_lines_one_zone_HBO_alt %>%
          st_cast("POINT") %>%
          st_coordinates() %>%
          .[!c(TRUE,FALSE),]) %>%
  cbind(desire_lines_one_zone_HBO_alt)

# HBW
desire_lines_coord_HBW_alt <- desire_lines_one_zone_HBW_alt %>%
  st_cast("POINT") %>%
  st_coordinates() %>%
  .[c(TRUE,FALSE),] %>%
  cbind(desire_lines_one_zone_HBW_alt %>%
          st_cast("POINT") %>%
          st_coordinates() %>%
          .[!c(TRUE,FALSE),]) %>%
  cbind(desire_lines_one_zone_HBW_alt)

# NHB
desire_lines_coord_NHB_alt <- desire_lines_one_zone_NHB_alt %>%
  st_cast("POINT") %>%
  st_coordinates() %>%
  .[c(TRUE,FALSE),] %>%
  cbind(desire_lines_one_zone_NHB_alt %>%
          st_cast("POINT") %>%
          st_coordinates() %>%
          .[!c(TRUE,FALSE),]) %>%
  cbind(desire_lines_one_zone_NHB_alt)

# All
desire_lines_coord_all_alt <- desire_lines_one_zone_all_alt %>%
  st_cast("POINT") %>%
  st_coordinates() %>%
  .[c(TRUE,FALSE),] %>%
  cbind(desire_lines_one_zone_all_alt %>%
          st_cast("POINT") %>%
          st_coordinates() %>%
          .[!c(TRUE,FALSE),]) %>%
  cbind(desire_lines_one_zone_all_alt)



ggplot(desire_lines_one_zone_HBO_alt) +
  # annotation_map_tile(type = "cartolight", zoom = 11, progress = "none") +
  geom_sf(data = zone_boundaries, fill = "grey35", color = alpha("white", .1), size = .1) +
  geom_sf(color = NA) +
  geom_curve(data = desire_lines_coord_HBO_alt,
             aes(x = X, y = Y, xend = X.1, yend = Y.1, 
                 alpha = flow,
                 color = flow), curvature = .2) +
  scale_color_viridis_c(option = "C",
                        name = "Trip Flow") +
  scale_alpha(guide = 'none') +
  theme_void()

#zoom in on downtown for more clarity
left_side <- st_bbox(zone_boundaries %>% filter(GEOID == HBO_big_attraction_alt))$xmin - .01
right_side <- st_bbox(zone_boundaries %>% filter(GEOID == HBO_big_attraction_alt))$xmax + .01
top_side <- st_bbox(zone_boundaries %>% filter(GEOID == HBO_big_attraction_alt))$ymax
bottom_side <- st_bbox(zone_boundaries %>% filter(GEOID == HBO_big_attraction_alt))$ymin

ggplot(desire_lines_one_zone_HBO_alt) +
  geom_sf(data = zone_boundaries, fill = "grey35", color = alpha("white", .1), size = .1) +
  geom_sf(color = NA) +
  geom_curve(data = desire_lines_coord_HBO_alt,
             aes(x = X, y = Y, xend = X.1, yend = Y.1, 
                 alpha = flow,
                 color = flow), curvature = .2) +
  coord_sf(xlim = c(left_side, right_side),
           ylim = c(bottom_side, top_side)) +
  scale_color_viridis_c(option = "C",
                        name = "Trip Flow",
                        # labels = c("3,000","6,000","9,000")
                        ) +
  scale_alpha(guide = 'none') +
  theme_void()



```

# Differences Between Existing and Alternative

```{r}

desire_lines_coord_all <- desire_lines_coord_all %>%
  as.data.frame() %>%
  full_join(desire_lines_coord_all_alt %>%
              as.data.frame(),
            by = c("X", "Y", "X.1", "Y.1", "o_id", "d_id", "geometry")) %>%
  mutate(hbo_flow_dif = (hbo_flow_alt - hbo_flow) / hbo_flow,
         hbw_flow_dif = (hbw_flow_alt - hbw_flow) / hbw_flow,
         nhb_flow_dif = (nhb_flow_alt - nhb_flow) / nhb_flow) %>%
  st_as_sf()

# Plot % difference in HBO trips
ggplot(desire_lines_coord_all[order(-desire_lines_coord_all$hbo_flow_dif),]) +
  geom_sf(data = zone_boundaries, color = alpha("black", .3), fill = "grey35") +
  geom_curve(aes(x = X, y = Y, xend = X.1, yend = Y.1,
                 color = hbo_flow_dif*100), curvature = .2, alpha = ifelse(desire_lines_coord_all$hbo_flow_dif > 0 &
                                                                             desire_lines_coord_all$hbo_flow_dif < .4,
                                                                           .1, 1)) +
  # coord_sf(xlim = c(left_side, right_side),
  #          ylim = c(bottom_side, top_side)) +
  scale_color_gradient2(low = "red", mid = "grey60", high = "limegreen",
                        name = "% HBO Flow Trip Difference\nExisting to Alternative") +
  # scale_alpha(guide = 'none') +
  theme_void()


# Plot % difference in HBW trips
ggplot(desire_lines_coord_all[order(-desire_lines_coord_all$hbw_flow_dif),]) +
  geom_sf(data = zone_boundaries, color = alpha("black", .3), fill = "grey35") +
  geom_curve(aes(x = X, y = Y, xend = X.1, yend = Y.1,
                 color = hbw_flow_dif*100), curvature = .2, alpha = ifelse(desire_lines_coord_all$hbw_flow_dif > -.1, .2, 1)) +
  # coord_sf(xlim = c(left_side, right_side),
  #          ylim = c(bottom_side, top_side)) +
  scale_color_gradient2(low = "red", mid = "grey60", high = "limegreen",
                        name = "% HBW Flow Trip Difference\nExisting to Alternative") +
  # scale_alpha(guide = 'none') +
  theme_void()


# Plot % difference in NHB trips
ggplot(desire_lines_coord_all[order(-desire_lines_coord_all$nhb_flow_dif),]) +
  geom_sf(data = zone_boundaries, color = alpha("black", .3), fill = "grey35") +
  geom_curve(aes(x = X, y = Y, xend = X.1, yend = Y.1,
                 color = nhb_flow_dif*100), curvature = .2, alpha = ifelse(desire_lines_coord_all$hbo_flow_dif > 0 &
                                                                             desire_lines_coord_all$hbo_flow_dif < .4,
                                                                           .1, 1)) +
  # coord_sf(xlim = c(left_side, right_side),
  #          ylim = c(bottom_side, top_side)) +
  scale_color_gradient2(low = "red", mid = "grey60", high = "limegreen",
                        name = "% NHB Flow Trip Difference\nExisting to Alternative") +
  # scale_alpha(guide = 'none') +
  theme_void()





```








