---
title: "assignment 7"
author: "Megan Willis-Jackson, Claire Wang, Ignacio Lafuente"
date: "3/3/2022"
output: html_document
---

```{r, message=F, echo=F, warning=F}
library(here)
library(tidyverse)
library(survey)
library(srvyr)
library(naniar)
library(jtools)
library(knitr)
```


#set up NHTS data
```{r, message=F, echo=F, warning=F}
temp <- tempfile()
download.file("https://nhts.ornl.gov/assets/2016/download/csv.zip", temp)

trips <- read_csv(unz(temp, "trippub.csv"))
person <- read_csv(unz(temp, "perpub.csv"))
hhs <- read_csv(unz(temp, "hhpub.csv"))

unlink(temp)
```


#select household-level variables
```{r, message=F, echo=F, warning=F}
sanjose_hhs <- hhs %>%
  filter(HH_CBSA == "41940") %>%
  mutate(inc_quint_ = case_when(HHFAMINC == "01" ~ "1st",
                               HHFAMINC == "02" ~ "1st",
                               HHFAMINC == "03" ~ "1st",
                               HHFAMINC == "04" ~ "2nd",
                               HHFAMINC == "05" ~ "2nd",
                               HHFAMINC == "06" ~ "3rd",
                               HHFAMINC == "07" ~ "4th",
                               HHFAMINC == "08" ~ "5th",
                               HHFAMINC == "09" ~ "5th",
                               HHFAMINC == "10" ~ "5th",
                               HHFAMINC == "11" ~ "5th",
                               TRUE ~ "NA")) %>%
  mutate(size_ = case_when(HHSIZE == 1 ~ "one",
                              HHSIZE == 2 ~ "two",
                              HHSIZE == 3 ~ "three",
                              TRUE ~ "four_plus")) %>%
  mutate(zero_veh_ = (HHVEHCNT == 0)) %>%
  replace_with_na(list(inc_quint_ = "NA")) %>%
  select(HOUSEID, zero_veh_, size_, inc_quint_, WRKCOUNT, WTHHFIN) # adding no. of workers at the hh

```

#select person-level variables
```{r, message=F, echo=F, warning=F}
# workers who work from home, have the option to do so, or have effectively worked from home more than one day during that month, will be considered.

# the data will be aggregated at the hh level and contrasted with the number of total workers there. The variable will measure how generation of trips is affected as WFH becomes more relevant as a % at the household level.

# extrapolation, although myopic (from disaggregated data to aggregated data), will apply its coefficients to the % of WFH at each zone.

sanjose_pers <- person %>%
  filter(HH_CBSA == "41940") %>%
  mutate(WFH_exh = case_when(WRK_HOME == '01' | WKRMHM == '01' | WKFMHMXX > 0 ~ 1,
         TRUE ~ 0)) %>%
  select(HOUSEID, WFH_exh)
```


#select trip-level variables
```{r, message=F, echo=F, warning=F}

# we decided to still count HBW trips for WFH individuals because it does not necessarily means never visiting the office or making trips for work purposes.

# moreover, if our assumptions are correct, this could be measured and prove a smaller generation.

trips_by_purpose <- trips %>% 
  filter(HH_CBSA == "41940") %>%
  select(HOUSEID, WHYFROM, WHYTO) %>%
  mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
                                WHYTO == "02" ~ TRUE,
                                WHYFROM == "01" ~ TRUE,
                                WHYFROM == "02" ~ TRUE,
                                TRUE ~ FALSE)) %>%
  mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
  mutate(purpose = case_when(home_based & work ~ "HBW",
                            home_based ~ "HBO",
                            TRUE ~ "NHB")) %>%
  group_by(HOUSEID, purpose) %>%
  summarize(n = n()) %>%
  pivot_wider(names_from = "purpose", values_from = "n") 

sanjose_pers_agg <- sanjose_pers %>%
  group_by(HOUSEID) %>%
  summarize(WFH_exh = sum(WFH_exh))

# #differentiate wfh household household trips from other home-based trip
# trips_by_purpose_Claire <- trips %>% 
#   filter(HH_CBSA == "41940") %>%
#   select(HOUSEID, WHYFROM, WHYTO) %>%
#   mutate(home_based = case_when(WHYTO == "01" ~ TRUE,
#                                 WHYFROM == "01" ~ TRUE,
#                                 TRUE ~ FALSE)) %>%
#   mutate(work = ifelse(WHYTO == "03" | WHYFROM == "03", TRUE, FALSE)) %>%
#   mutate(wfh = ifelse(WHYTO == "02" | WHYFROM == "02", TRUE, FALSE)) %>%
#   mutate(purpose = case_when(home_based & work ~ "HBW",
#                             home_based ~ "HBO",
#                             wfh ~ "WFHB",
#                             TRUE ~ "NHB")) %>%
#   group_by(HOUSEID, purpose) %>%
#   summarize(n = n()) %>%
#   pivot_wider(names_from = "purpose", values_from = "n") 

```

#join trip and person data to hh data

```{r, message=F, echo=F, warning=F}

# joins using households as rows

hh_pers_trips <- left_join(sanjose_hhs, trips_by_purpose) %>%
  replace_na(list(HBW = 0,
                  HBO = 0,
                  NHB = 0)) %>%
  left_join(sanjose_pers_agg) %>%
  mutate(wfh_ratio = WFH_exh / WRKCOUNT) %>%
  replace_na(list(wfh_ratio = 0)) %>%
  mutate(wfh_int_ = case_when(wfh_ratio <= 0.5 & wfh_ratio > 0.25 ~ 'mid',
                              wfh_ratio > 0.5 ~ 'high',
                             TRUE ~ 'low'))

# let a low-wfh intensity be the reference for regs

hh_pers_trips$wfh_int_ <- relevel(factor(hh_pers_trips$wfh_int_), ref="low")
```

#CREATE SURVEY OBJECT

```{r, message=F, echo=F, warning=F}

svy_trips <- hh_pers_trips %>%
  as_survey(weights = WTHHFIN)

```

#Estimate household-level regression models

```{r, message=F, echo=F, warning=F}
HBO_model1 <- svyglm(HBO ~ zero_veh_ + size_ + inc_quint_ + wfh_int_, svy_trips)

export_summs(HBO_model1, 
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = "Full model")

#remote non-statistically significant vars (zero_veh_hh & income quantiles, wfh preserved)

HBO_model2 <- svyglm(HBO ~ size_ + wfh_int_, svy_trips)

export_summs(HBO_model1, HBO_model2,
             error_pos = "right",
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model"))
```

```{r, message=F, echo=F, warning=F}
HBW_model1 <- svyglm(HBW ~ zero_veh_ + size_ + inc_quint_ + wfh_int_, svy_trips)

export_summs(HBW_model1, 
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = "Full model")

#remote non-statistically significant vars (zero_veh_hh & income quantiles, wfh preserved)

HBW_model2 <- svyglm(HBW ~ size_ + wfh_int_, svy_trips)

export_summs(HBW_model1, HBW_model2,
             error_pos = "right",
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model"))

```

```{r, message=F, echo=F, warning=F}
NHB_model1 <- svyglm(NHB ~ zero_veh_ + size_ + inc_quint_ + wfh_int_, svy_trips)

export_summs(NHB_model1, 
             error_pos = "right", 
             error_format = "(p = {p.value})",
             model.names = "Full model")

#remote non-statistically significant vars (zero_veh_hh & income quantiles, wfh preserved)

NHB_model2 <- svyglm(NHB ~ size_ + wfh_int_, svy_trips)

export_summs(NHB_model1, NHB_model2,
             error_pos = "right",
             error_format = "(p = {p.value})",
             model.names = c("Full model", "Reduced model"))

```
#Apply trip production model to existing zonal data
```{r, message=F, echo=F, warning=F}
existing_zones <- here("existing",
                       "data",
                       "existing_sanjose.csv") %>%
  read_csv() %>%
  mutate(wfh_ratio = total_wfhE / total_workerE) %>%
  mutate(wfh_int = case_when(wfh_ratio <= 0.05 ~ 'low',
                             wfh_ratio <= 0.075 & wfh_ratio > 0.05 ~ 'mid',
                             TRUE ~ 'high'))
# (see brackets justification below)

existing_zones <- existing_zones %>%
  mutate(hbo_prod = total_hhsE * HBO_model2$coefficients["(Intercept)"] +
           hh_1personE * HBO_model2$coefficients["size_one"] +
           hh_2personE * HBO_model2$coefficients["size_two"] +
           hh_3personE * HBO_model2$coefficients["size_three"] +
           total_hhsE * if (existing_zones$wfh_int == 'mid') {HBO_model2$coefficients["wfh_int_mid"]}
         else if (existing_zones$wfh_int == 'high') {HBO_model2$coefficients["wfh_int_high"]}
         else {0}) %>%
  mutate(hbw_prod = total_hhsE * HBW_model2$coefficients["(Intercept)"] +
           hh_1personE * HBW_model2$coefficients["size_one"] +
           hh_2personE * HBW_model2$coefficients["size_two"] +
           hh_3personE * HBW_model2$coefficients["size_three"] +
           total_hhsE * if (existing_zones$wfh_int == 'mid') {HBW_model2$coefficients["wfh_int_mid"]}
         else if (existing_zones$wfh_int == 'high') {HBW_model2$coefficients["wfh_int_high"]}
         else {0}) %>%
  mutate(nhb_prod = total_hhsE * NHB_model2$coefficients["(Intercept)"] +
           hh_1personE * NHB_model2$coefficients["size_one"] +
           hh_2personE * NHB_model2$coefficients["size_two"] +
           hh_3personE * NHB_model2$coefficients["size_three"] +
           total_hhsE * if (existing_zones$wfh_int == 'mid') {NHB_model2$coefficients["wfh_int_mid"]}
         else if (existing_zones$wfh_int == 'high') {NHB_model2$coefficients["wfh_int_high"]}
         else {0})

# histograms showing distributions for wfh_ratio at the census tract- and household-level, respectively. Approximate deciles were used to trace categorical parallelisms between the aggregated and disaggregated data.

hist(existing_zones$wfh_ratio)
hist(hh_pers_trips$wfh_ratio)
quantile(existing_zones$wfh_ratio, probs = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1))
quantile(hh_pers_trips$wfh_ratio, probs = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1))

```

#Apply trip production model to alternative zonal data

```{r, message=F, echo=F, warning=F}
alternative_zones <- here("alternative",
                       "data",
                       "alternative_sanjose.csv") %>%
  read_csv() %>%
  mutate(wfh_ratio = alt_wfh / total_workerE) %>%
  mutate(wfh_int = case_when(wfh_ratio <= 0.05 ~ 'low',
                             wfh_ratio <= 0.075 & wfh_ratio > 0.05 ~ 'mid',
                             TRUE ~ 'high'))

alternative_zones <- alternative_zones %>%
  mutate(hbo_prod = total_hhsE * HBO_model2$coefficients["(Intercept)"] +
           hh_1personE * HBO_model2$coefficients["size_one"] +
           hh_2personE * HBO_model2$coefficients["size_two"] +
           hh_3personE * HBO_model2$coefficients["size_three"] +
           total_hhsE * if (alternative_zones$wfh_int == 'mid') {HBO_model2$coefficients["wfh_int_mid"]}
         else if (alternative_zones$wfh_int == 'high') {HBO_model2$coefficients["wfh_int_high"]}
         else {0}) %>%
  mutate(hbw_prod = total_hhsE * HBW_model2$coefficients["(Intercept)"] +
           hh_1personE * HBW_model2$coefficients["size_one"] +
           hh_2personE * HBW_model2$coefficients["size_two"] +
           hh_3personE * HBW_model2$coefficients["size_three"] +
           total_hhsE * if (alternative_zones$wfh_int == 'mid') {HBW_model2$coefficients["wfh_int_mid"]}
         else if (alternative_zones$wfh_int == 'high') {HBW_model2$coefficients["wfh_int_high"]}
         else {0}) %>%
  mutate(nhb_prod = total_hhsE * NHB_model2$coefficients["(Intercept)"] +
           hh_1personE * NHB_model2$coefficients["size_one"] +
           hh_2personE * NHB_model2$coefficients["size_two"] +
           hh_3personE * NHB_model2$coefficients["size_three"] +
           total_hhsE * if (alternative_zones$wfh_int == 'mid') {NHB_model2$coefficients["wfh_int_mid"]}
         else if (alternative_zones$wfh_int == 'high') {NHB_model2$coefficients["wfh_int_high"]}
         else {0})

```

#Trip attraction

```{r, message=F, echo=F, warning=F}
# existing conditions

existing_zones <- existing_zones %>%
  replace_na(list(basic_jobs = 0,
                  retail_jobs = 0, 
                  service_jobs = 0)) %>%
  mutate(hbo_attr = 0.7 * total_hhsE +
                    0.7 * basic_jobs +
                    8.4 * retail_jobs +
                    3.5 * service_jobs,
         hbw_attr = 1.2 * total_jobs,
         nhb_attr = 1.4 * total_hhsE +
                    6.9 * retail_jobs +
                    0.9 * service_jobs)

# alternative conditions

alternative_zones <- alternative_zones %>%
  replace_na(list(basic_jobs = 0,
                  retail_jobs = 0, 
                  service_jobs = 0)) %>%
  mutate(hbo_attr = 0.7 * total_hhsE +
                    0.7 * basic_jobs +
                    8.4 * retail_jobs +
                    3.5 * alt_service_jobs,
         hbw_attr = 1.2 * (total_jobs - jobs_change),
         nhb_attr = 1.4 * total_hhsE +
                    6.9 * retail_jobs +
                    0.9 * alt_service_jobs)

```

#Balancing

```{r, message=F, echo=F, warning=F}

# checking for initial differences

trip_end_summary <- tibble(Purpose = c("ex_HBO", "ex_HBW", "ex_NHB", "alt_HBO", "alt_HBW", "alt_NHB"),
                           Productions = c(sum(existing_zones$hbo_prod), sum(existing_zones$hbw_prod), sum(existing_zones$nhb_prod), sum(alternative_zones$hbo_prod), sum(alternative_zones$hbw_prod), sum(alternative_zones$nhb_prod)),
                           Attractions = c(sum(existing_zones$hbo_attr), sum(existing_zones$hbw_attr), sum(existing_zones$nhb_attr), sum(alternative_zones$hbo_attr), sum(alternative_zones$hbw_attr), sum(alternative_zones$nhb_attr)),
                           Difference = c(sum(existing_zones$hbo_attr) -                    sum(existing_zones$hbo_prod),
                                          sum(existing_zones$hbw_attr) - sum(existing_zones$hbw_prod),
                                          sum(existing_zones$nhb_attr) - sum(existing_zones$nhb_prod),
                                          sum(alternative_zones$hbo_attr) -                 sum(alternative_zones$hbo_prod),
                                          sum(alternative_zones$hbw_attr) - sum(alternative_zones$hbw_prod),
                                          sum(alternative_zones$nhb_attr) - sum(alternative_zones$nhb_prod)))

kable(trip_end_summary, format.args = list(big.mark = ","))

# balancing existing conditions

existing_zones <- existing_zones %>%
  mutate(hbo_attr_bal = hbo_attr * 
           sum(existing_zones$hbo_prod) / sum(existing_zones$hbo_attr),
         hbw_attr_bal = hbw_attr * 
           sum(existing_zones$hbw_prod) / sum(existing_zones$hbw_attr),
         nhb_attr_bal = nhb_attr * 
           sum(existing_zones$nhb_prod) / sum(existing_zones$nhb_attr))

# balancing alternative conditions

alternative_zones <- alternative_zones %>%
  mutate(hbo_attr_bal = hbo_attr * 
           sum(alternative_zones$hbo_prod) / sum(alternative_zones$hbo_attr),
         hbw_attr_bal = hbw_attr * 
           sum(alternative_zones$hbw_prod) / sum(alternative_zones$hbw_attr),
         nhb_attr_bal = nhb_attr * 
           sum(alternative_zones$nhb_prod) / sum(alternative_zones$nhb_attr))

#checking after balancing

trip_end_summary_bal <- tibble(Purpose = c("ex_HBO", "ex_HBW", "ex_NHB", "alt_HBO", "alt_HBW", "alt_NHB"),
                           Productions = c(sum(existing_zones$hbo_prod), sum(existing_zones$hbw_prod), sum(existing_zones$nhb_prod), sum(alternative_zones$hbo_prod), sum(alternative_zones$hbw_prod), sum(alternative_zones$nhb_prod)),
                           Attractions = c(sum(existing_zones$hbo_attr_bal), sum(existing_zones$hbw_attr_bal), sum(existing_zones$nhb_attr_bal), sum(alternative_zones$hbo_attr_bal), sum(alternative_zones$hbw_attr_bal), sum(alternative_zones$nhb_attr_bal)),
                           Difference = c(sum(existing_zones$hbo_attr_bal) -                    sum(existing_zones$hbo_prod),
                                          sum(existing_zones$hbw_attr_bal) - sum(existing_zones$hbw_prod),
                                          sum(existing_zones$nhb_attr_bal) - sum(existing_zones$nhb_prod),
                                          sum(alternative_zones$hbo_attr_bal) -                 sum(alternative_zones$hbo_prod),
                                          sum(alternative_zones$hbw_attr_bal) - sum(alternative_zones$hbw_prod),
                                          sum(alternative_zones$nhb_attr_bal) - sum(alternative_zones$nhb_prod)))

kable(trip_end_summary_bal, format.args = list(big.mark = ","))
```

Interpretation of results so far: more working from home would increase the number of non-work related trips from home (HBO), which resonates with the flexibility these workers gain. Home-based trips to work (HBW) decrease, which is the expected outcome by definition, following the logical trend. Non-home based trips (NHB) should a priori decrease too, given that the typical example would be a trip generated at work (e.g., lunch). However, NHB trips increase in a large proportion, and possibly only sophisticated patterns of trip chaining could explain that intuitively. Overall, the MSA would incorporate 1.1 M trips, posing a question on how many benefits could be expected from WFH if 'removing commuters from the streets' in turn propagates their generation of other trips more than proportionately.

#create some chloropleth maps and/or dot-density maps

# save data
```{r, message=F, echo=F, warning=F}
write_csv(existing_zones, file = here("existing", 
                                      "data",
                                      "existing_sanjose.csv"))
write_csv(alternative_zones, file = here("alternative", 
                                      "data",
                                      "alternative_sanjose.csv"))
```

