---
title: "Assignment02"
author: "Megan Willis-Jackson, Claire Wang, Ignacio Lafuente"
date: "2/3/2022"
output: html_document
---

```{r libraries, echo=F, message=F, warning=F}
library(tidyverse)
library(tidycensus)
library(ggplot2)
library(sf)
library(units)
library(tigris)
library(magrittr)
library(dplyr)
library(RColorBrewer)
library(viridis) 
library(GGally)
library(vtable)
library(extrafont)

```

# Load Variables

```{r, message = F, echo = F, error = F}
sanjose_msa <- core_based_statistical_areas() %>% 
  filter(GEOID == "41940")

# view(load_variables(2019, "acs5"))

#specify ACS variables for inclusion
vars <- c(med_income = "B06011_001",
          total_hhs = 'B08203_001',
          no_veh_hh = 'B08203_002',
          one_veh_hh = 'B08201_003',
          two_veh_hh = 'B08201_004',
          three_veh_hh = 'B08201_005',
          fourmore_veh_hh = 'B08201_006',
          no_worker_hh = 'B08202_002',
          one_worker_hh = 'B08202_003',
          two_worker_hh = 'B08202_004',
          threemore_worker_hh = 'B08202_005',
          hh_1person = 'B08201_007',
          hh_2person = 'B08201_013',
          hh_3person = 'B08201_019',
          hh_4person_plus = 'B08201_025',
          inc_lt_10k = 'B19001_002',
          inc_btw_10k_15k = 'B19001_003',
          inc_btw_15k_20k = 'B19001_004',
          inc_btw_20k_25k = 'B19001_005',
          inc_btw_25k_30k = 'B19001_006',
          inc_btw_30k_35k = 'B19001_007',
          inc_btw_35k_40k = 'B19001_008',
          inc_btw_40k_45k = 'B19001_009',
          inc_btw_45k_50k = 'B19001_010',
          inc_btw_50k_60k = 'B19001_011',
          inc_btw_60k_75k = 'B19001_012',
          inc_btw_75k_100k = 'B19001_013',
          inc_btw_100k_125k = 'B19001_014',
          inc_btw_125k_150k = 'B19001_015',
          inc_btw_150k_200k = 'B19001_016',
          inc_gt_200k = 'B19001_017',
          total_worker = 'B08301_001',
          total_wfh = 'B08301_021') 

#save ACS variables
census <- get_acs(geography = "tract",
                  state = "CA",
                  variables = vars,
                  output = "wide",
                  geometry = T) %>%
  st_filter(sanjose_msa)

ca_tracts <- get_acs(geography = "tract",
                  state = "CA",
                  variables = "B06011_001",
                  output = "wide",
                  geometry = T)

#specify location as San Jose MSA  
sj_centroids <- st_centroid(ca_tracts) %>%
  st_filter(sanjose_msa) 

tracts <- ca_tracts %>%
  st_filter(sj_centroids)
  
#download employment data
lehd_blocks_jobs <- read_csv("https://lehd.ces.census.gov/data/lodes/LODES7/ca/wac/ca_wac_S000_JT00_2019.csv.gz", 
                             show_col_types = F) %>%
  rename(total_jobs = C000) %>%
  mutate(basic_jobs = CNS01+CNS02+CNS03+CNS04+CNS05+CNS06+CNS08+CNS09) %>%
  rename(retail_jobs = CNS07) %>%
  mutate(service_jobs = total_jobs - basic_jobs - retail_jobs) %>%
  select(w_geocode, total_jobs, basic_jobs, retail_jobs, service_jobs)


#collapse LEHD variables to census tracts
lehd_tracts_jobs <- lehd_blocks_jobs %>%
  mutate(w_geocode = as.character(w_geocode)) %>%
  mutate(GEOID = substr(w_geocode, 1, 11)) %>%
  select(-w_geocode) %>%
  group_by(GEOID) %>%
  summarise(across(everything(), ~sum(.))) 

#join ACS data with LEHD data
census <- census %>%
  left_join(lehd_tracts_jobs, by = "GEOID") %>%
  select(-contains("M", ignore.case = F))


```

```{r}
# overall statistics for the MSA, focusing on the most vulnerable population
census <- census %>%
  mutate(ratio_no_veh = no_veh_hhE / total_hhsE,
         ratio_no_worker = no_worker_hhE / total_hhsE,
         ratio_lt_10k = inc_lt_10kE / total_hhsE,
         non_wfh = total_workerE - total_wfhE)

sumtable(census, vars = c('ratio_no_veh', 'ratio_no_worker', 'ratio_lt_10k'), add.median = TRUE, title = 'Summary Statistics - Vulnerability Ratios', col.width = c(10))
```

```{r}
# distributions for income and vehicle ownership

hist_y <- ggplot(census) +
  geom_histogram(aes(x = census$med_income),
                 fill = "dodgerblue3") +
  labs(y = "Count", x = "Median Income")

hist_y

veh <- data.frame(veh_own = c("0","1","2","3","4+"), count = c(sum(census$no_veh_hhE) / sum(census$total_hhsE), sum(census$one_veh_hhE) / sum(census$total_hhsE), sum(census$two_veh_hhE) / sum(census$total_hhsE), sum(census$three_veh_hhE) / sum(census$total_hhsE), sum(census$fourmore_veh_hhE) / sum(census$total_hhsE)))

hist_veh <- ggplot(veh, aes(x = veh_own, y = count)) +
  geom_bar(stat = "identity", fill = "firebrick3") +
  labs(y = "Density", x = "Vehicles Owned by Household")

hist_veh
```


# Create Alternative Dataset

The main simplifying assumption used in creating the alternative dataset is
that all jobs within service sector are able to be performed remotely,
and that people working in those positions will decrease their
days spent in the office by 40% (i.e., the number of people working
from home will increase by 40%). Additionally, the number of jobs
in the service sector will decrease by 40%. These jobs are not
being removed from the economy, but for the purposes of the subsequent
transportation demand model, they will not be considered as jobs
that require commuters.

The main assumption here is that the industries that we have classified
as "WFH" will experience a 40% decrease in work days commuting from the 
baseline, i.e. from 5 days per week to 3 days per week.


```{r}
census_alt <- census %>%
  mutate(wfh_percent = .4,
         alt_service_jobs = service_jobs * (1 - wfh_percent),
         jobs_change = service_jobs * wfh_percent,
         commuter_change = non_wfh * wfh_percent,
         alt_non_wfh = non_wfh - commuter_change,
         alt_wfh = total_wfhE + commuter_change)


```

# Generate Graphics

```{r}

#color palettes
map_color<- brewer.pal(5,"BuPu")
map_color2<-brewer.pal(5,"PRGn")
map_color3<-brewer.pal(5,"YlOrRd")
map_color4<-brewer.pal(5,"Oranges")
map_color5<-brewer.pal(5,"PuRd")
map_color6<-brewer.pal(5,"Greens")
map_color7<-brewer.pal(5,"Reds")

#specify text elements to help standardize
font <- "Trebuchet MS"
text_size <- 12
axis_title_size <- 14
axis_text_size <- 12
geom_text_size <- 4
options(scipen = 999)



#create a function to help standardize maps. Can be updated as needed
chlor_map <- function(df, variable, pal = NULL, title = NULL) {
  ggplot(df) +
    geom_sf(aes(fill = .data[[variable]]), color = NA) +
  scale_fill_gradientn(name = title,
                       colors = pal) +
    theme_void() +
    theme(panel.grid.minor = element_blank(), 
          text = element_text(family = font, size = text_size))

}

wfhemp_change_map <- chlor_map(census_alt, "commuter_change", pal = map_color,
                             title = "Commuters\nTransitioned\nto WFH")


alt_wfhemp_map <- chlor_map(census_alt, "alt_wfh", pal = map_color,
                            title = "Primarily WFH in\nAlternative Scenario")


wfhjobs_change_map <- chlor_map(census_alt, "jobs_change", pal = map_color,
                                title = "Jobs Transitioned\nto WFH in\nAlternative Scenario")


ggsave("graphics/employee_commuting_change.jpeg", wfhemp_change_map, device = "jpeg")
ggsave("graphics/job_commuting_change.jpeg", wfhjobs_change_map, device = "jpeg")
ggsave("graphics/employee_wfh_map.jpeg", alt_wfhemp_map, device = "jpeg")



```


```{r}


overall <- ggplot(census_alt, 
       aes(x = alt_wfh/total_workerE,#x = alt_wfh_emp/alt_total_emp,
           y = med_incomeE/1000,
           size = no_veh_hhE/total_hhsE)) +
  geom_point(alpha = 0.5) +
  scale_size_continuous(name = "Proportion of households\nwith no vehicles") +
  scale_x_continuous(name = "Employees able to WFH / Total employees",
                     breaks = seq(0, 1, by = 0.1),
                     labels = paste(seq(0, 100, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Median Income (in thousands)",
                     breaks = seq(0, 120, by = 10)) +
  theme_minimal() +
  theme(legend.key.size = unit(0.1, "cm"))

overall
  
newdf <- census_alt[,c("med_incomeE","alt_wfh","alt_service_jobs")] %>% 
  st_set_geometry(NULL)

corr <- ggcorr(newdf,palette = "RdBu",label=TRUE)

income_map <- chlor_map(census, "med_incomeE", pal = map_color,
                        title = "Median Income")

jobs_map <- chlor_map(census, "total_jobs", pal = map_color,
                      title = "Existing Total Jobs")

employees_map <- chlor_map(census, "total_workerE", pal = map_color,
                           title = "Existing Total Employees")

noveh_map <- chlor_map(census, "ratio_no_veh", pal = map_color,
                       title = "Ratio Households with\nNo Vehicle to\nOne or More")

noworker_map <- chlor_map(census, "ratio_no_worker", pal = map_color,
                          title = "Ratio Households with\nNo Workers to\nOne or More")

lowinc_map <- chlor_map(census, "ratio_lt_10k", pal = map_color,
                        title = "Ratio Households <$10k Income")



ggsave("graphics/dotplot.jpeg", overall, device = "jpeg")
ggsave("graphics/correlation.jpeg", corr, device = "jpeg")
ggsave("graphics/income_map.jpeg", income_map, device = "jpeg")
ggsave("graphics/jobs_map.jpeg", jobs_map, device = "jpeg")
ggsave("graphics/employees_map.jpeg", employees_map, device = "jpeg")
ggsave("graphics/novehicles_map.jpeg", noveh_map, device = "jpeg")
ggsave("graphics/noworkers_map.jpeg", noworker_map, device = "jpeg")
ggsave("graphics/lowincome_map.jpeg", lowinc_map, device = "jpeg")



```

# Save Datasets

```{r}
existing <- write_csv(census, "existing/data/existing_sanjose.csv", append = F)

zones_sanjose <- write_sf(tracts, "zones/zones_sanjose2.geojson")

alternative <- write_csv(census_alt, "alternative/data/alternative_sanjose.csv", append = F)

```









